---
title: "Question 2"
author: "gowtham_1p23mb010"
date: "2025-02-02"
output: html_document
--- 
 Question No: 02
 
Using the "Marketing Customer Value Analysis" dataset, complete the following tasks with proper analysis and interpretation: 
 
 
i.	Load the dataset and explore its structure using basic commands.

```{r}
library(ggplot2)
```

```{r}
df=read.csv("D:/Data Science for Marketing-I/data/WA_Fn-UseC_-Marketing-Customer-Value-Analysis.csv")
```
```{r}
head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```

```{r}
dim(df)
```
Interpertation: Basic commands provide an overview of the dataset.

ii.	Create a new column named "Engaged" by transforming the categorical values in the "Response" variable into numerical values. Why is this transformation important?


```{r}
df$Engaged <- ifelse(df$Response == "Yes", 1, 0)

```

Interpertation: Converts "Yes/No" responses into 1 and 0 for numerical analysis.

iii.	Calculate and interpret the Engagement Rate. How is it computed, and what does it indicate about the customer responses?

```{r}
print(sum(df$Engaged)/nrow(df)*100)
```
Interpertation: Engagement rate shows the percentage of customers who responded positively.

iv.	Analyze engagement rate by "Renew Offer Type" and "Sales Channel":

```{r}
aggregate(Engaged ~ Renew.Offer.Type + Sales.Channel, data = df, mean)
```

Interpertation: Groups data by "Renew Offer Type" and "Sales Channel" to analyze engagement patterns.


v.	Use a pivot table to summarize engagement by "Renew Offer Type" and visualize the results using both bar and pie charts. Why are these visualizations helpful in understanding customer engagement patterns? 

```{r}
#Bar chart
ggplot(df, aes(x = Renew.Offer.Type, y = Engaged)) +
  stat_summary(fun = mean, geom = "bar", fill = "skyblue") +
  ggtitle("Engagement by Renew Offer Type")+
   theme_minimal() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
# Pie chart
engagement_summary <- aggregate(Engaged ~ Renew.Offer.Type, data = df, mean)
pie(engagement_summary$Engaged, labels = engagement_summary$Renew.Offer.Type, 
    main = "Engagement by Renew Offer Type", col = rainbow(length(engagement_summary$Engaged)))
```

 vi.	Explain the purpose of regression analysis in this context. Describe how you would approach regression using
 
 
**Purpose of Regression Analysis:** 
Regression analysis is used to understand and quantify the relationship between variables, enabling the prediction of an outcome variable (e.g., engagement) based on other influencing factors.
 
(i) continuous variables only

```{r}
# Continuous variables only
model_cont <- lm(Engaged ~ Income, data = df)
summary(model_cont)
```

Interpretation:
       **Continuous Variables:** Examines numerical predictors to assess their direct influence on the outcome. The coefficients indicate how much the target variable changes with a one-unit increase in the predictor. 

 (ii) categorical variables
 
```{r}
# Categorical variables
df$Renew.Offer.Type <- as.factor(df$Renew.Offer.Type)
model_cat <- glm(Engaged ~ Renew.Offer.Type, family = binomial, data = df)
summary(model_cat)
```
Interpretation: **Categorical Variables:** Utilizes dummy coding to evaluate the impact of different categories. The coefficients signify the variation from the baseline category.  


(iii) both continuous and categorical variables

```{r}
# Both types
model_both <- glm(Engaged ~ Income + Renew.Offer.Type, family = binomial, data = df)
summary(model_both)
```

Interpretation:**Combination of Both:** Integrates numerical and categorical variables for a comprehensive model. Continuous predictors show direct effects, while categorical variables are interpreted relative to a reference group.
























